---
title: "Assignment-2"
author: "eg"
date: "16/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("bnlearn")
#devtools::install_github("jtextor/dagitty/r")

#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")

#BiocManager::install("Rgraphviz")
```


Import useful libraries
```{r}
library(tidyverse)
library(dagitty)
library(lavaan)
library(stats)
library(bnlearn)
library(naniar)
library(ggplot2)
library(caret)
library(MLmetrics)
library(tree)
library(ROSE)
```



Read original data
```{r}
data_original <- read.csv(paste0("data\\HR-Employee-Attrition.csv"))
data_original <- as.data.frame(data_original)
```

#######################################################
Preprocessing


```{r}
colnames(data_original)[1] = 'Age'

data_original$Attrition <- as.numeric(ordered( data_original$Attrition,c("No","Yes")))
data_original$Attrition <- as.factor(as.numeric(ifelse(data_original$Attrition == 1, 0,1)))

data <- data_original %>% 
  mutate(TotalSatisfaction = rowMeans(cbind(EnvironmentSatisfaction, JobSatisfaction, RelationshipSatisfaction))) %>% 
  select(-c(Over18, StandardHours, EmployeeCount, EmployeeNumber, MonthlyRate, HourlyRate, DailyRate, JobInvolvement, EducationField, TrainingTimesLastYear,YearsWithCurrManager, Department, JobSatisfaction, EnvironmentSatisfaction, RelationshipSatisfaction, YearsInCurrentRole,  JobRole, PerformanceRating))



str(data)
```



```{r}
data$Age <- as.factor(cut(data$Age, c(17,29,40,50,60), include.lowest = T,ordered_result = T))

data$DistanceFromHome <- as.factor(cut(data$DistanceFromHome, c(0,5,15,30), include.lowest = T,ordered_result = T))

data$JobLevel <- as.factor(data$JobLevel)
data$JobLevel <- ordered(data$JobLevel, levels = c(1,2,3,4,5))

data$MonthlyIncome <- as.factor(cut(data$MonthlyIncome, c(0,3000,8000,12000,20000), include.lowest = T,ordered_result = T))

data$PercentSalaryHike <- as.factor(cut(data$PercentSalaryHike, c(11,13,17,25), include.lowest = T,ordered_result = T))

data$NumCompaniesWorked <- as.factor(cut(data$NumCompaniesWorked, c(0,1,4,9), include.lowest = T,ordered_result = T))

data$TotalWorkingYears <- as.factor(cut(data$TotalWorkingYears, c(0,5,10,20,40), include.lowest = T,ordered_result = T))

data$WorkLifeBalance <- as.factor(data$WorkLifeBalance)
data$WorkLifeBalance <- ordered(data$WorkLifeBalance, levels = c(1,2,3,4))

data$YearsAtCompany <- as.factor(cut(data$YearsAtCompany, c(0,5,10,20,40), include.lowest = T,ordered_result = T))

data$YearsSinceLastPromotion <- as.factor(cut(data$YearsSinceLastPromotion, c(0,1,3,5,10,15), include.lowest = T,ordered_result = T))

data$TotalSatisfaction <- as.factor(round(data$TotalSatisfaction,0))
data$TotalSatisfaction <- ordered(data$TotalSatisfaction, levels = c(1,2,3,4))

data$Education <- as.factor(data$Education)
data$Education <- ordered(data$Education, levels = c(1,2,3,4,5))

data$BusinessTravel <- as.factor(data$BusinessTravel)
data$BusinessTravel <- ordered(data$BusinessTravel, levels = c("Non-Travel","Travel_Rarely","Travel_Frequently"))

data$Gender <- as.factor(as.numeric(ordered( data$Gender,c("Female","Male"))))

data$MaritalStatus <-as.factor(as.integer( data$MaritalStatus=="Married" ))

data$StockOptionLevel <- as.factor(data$StockOptionLevel)
data$StockOptionLevel <-ordered(data$StockOptionLevel, levels = c(0,1,2,3))
```


```{r}
str(data)
```

```{r}
names <- c(colnames(data))
for (var in names) {
  print(var)
  print(levels(data[,var]))
}
```

We can see that we have 3 variables containing 5 different classes: Education, JobLevel, YearsSinceLastPromotion. We'll now try to reduce their number.

1) Education
We will divide this variable in three categories: 'College and below', 'Bachelor's and Master', 'Doctor'

```{r}
data$Education <-as.character(data$Education)
data$Education[data$Education== 1] <- "College and below"
data$Education[data$Education== 2] <- "College and below"
data$Education[data$Education == 3] <- "Bachelor's and Master"
data$Education[data$Education == 4] <- "Bachelor's and Master"
data$Education[data$Education==5] <- "Doctor"
data$Education <-as.factor(data$Education)

```

2) JobLevel
We will divide this variable in two categories: 'low', from 1 to 3 and 'high', 4 and 5.

```{r}
data$JobLevel <-as.character(data$JobLevel)
data$JobLevel[data$JobLevel== 1] <- "low"
data$JobLevel[data$JobLevel== 2] <- "low"
data$JobLevel[data$JobLevel == 3] <- "low"
data$JobLevel[data$JobLevel == 4] <- "high"
data$JobLevel[data$JobLevel==5] <- "high"
data$JobLevel <-as.factor(data$JobLevel)

```

3) yearsSinceLastPromotion
We will divide this variable in two categories: '0-1 years' and '2 or before'.

```{r}
data$YearsSinceLastPromotion <-as.character(data$YearsSinceLastPromotion)
data$YearsSinceLastPromotion[data$YearsSinceLastPromotion== 0] <- "this or last year"
data$YearsSinceLastPromotion[data$YearsSinceLastPromotion== 1] <- "this or last year"
data$YearsSinceLastPromotion[data$YearsSinceLastPromotion >= 2] <- "more than 2"
data$YearsSinceLastPromotion <-as.factor(data$YearsSinceLastPromotion)

```

############################################################################################
Constraint-based structure learning algorithms: PC algorithm 
and further preprocessing

```{r}
plot(pc.stable(data) )
```

Considering subset of variables:
```{r}
data_parents <- data %>% 
  select(c(Attrition, TotalSatisfaction, BusinessTravel, MonthlyIncome, MaritalStatus, OverTime, DistanceFromHome, Age, StockOptionLevel))

```

```{r}
plot(pc.stable(data_parents) )
```

We can see that also leaving only the Attrition partents with the highest estimate coefficient, we cannot achieve good results.
So, let's try to simplify more the variables:
```{r}
names <- c(colnames(data_parents))
for (var in names) {
  print(var)
  print(levels(data_parents[,var]))
}
```

```{r}
data_parents$TotalSatisfaction <-as.character(data_parents$TotalSatisfaction)
data_parents$TotalSatisfaction[data_parents$TotalSatisfaction== 1] <- "low"
data_parents$TotalSatisfaction[data_parents$TotalSatisfaction== 2] <- "low"
data_parents$TotalSatisfaction[data_parents$TotalSatisfaction == 3] <- "high"
data_parents$TotalSatisfaction[data_parents$TotalSatisfaction == 4] <- "high"
data_parents$TotalSatisfaction <-as.factor(data_parents$TotalSatisfaction)

```

```{r}
data_parents$MonthlyIncome <-as.character(data_parents$MonthlyIncome)
data_parents$MonthlyIncome[data_parents$MonthlyIncome== "[0,3e+03]"] <- "low"
data_parents$MonthlyIncome[data_parents$MonthlyIncome== "(3e+03,8e+03]"] <- "medium"
data_parents$MonthlyIncome[data_parents$MonthlyIncome == "(8e+03,1.2e+04]"] <- "medium"
data_parents$MonthlyIncome[data_parents$MonthlyIncome == "(1.2e+04,2e+04]"] <- "high"
data_parents$MonthlyIncome <-as.factor(data_parents$MonthlyIncome)

```

```{r}
data_parents$StockOptionLevel <-as.character(data_parents$StockOptionLevel)
data_parents$StockOptionLevel[data_parents$StockOptionLevel== 0] <- "No stock"
data_parents$StockOptionLevel[data_parents$StockOptionLevel== 1] <- "stock"
data_parents$StockOptionLevel[data_parents$StockOptionLevel == 2] <- "stock"
data_parents$StockOptionLevel[data_parents$StockOptionLevel == 3] <- "stock"
data_parents$StockOptionLevel <-as.factor(data_parents$StockOptionLevel)

```

```{r}
data_parents$BusinessTravel <-as.character(data_parents$BusinessTravel)
data_parents$BusinessTravel[data_parents$BusinessTravel== "Non-Travel"] <- "Non-Travel" 
data_parents$BusinessTravel[data_parents$BusinessTravel== "Travel_Rarely" ] <- "Travel"
data_parents$BusinessTravel[data_parents$BusinessTravel == "Travel_Frequently"] <- "Travel"
data_parents$BusinessTravel <-as.factor(data_parents$BusinessTravel)

```

```{r}
data_parents$DistanceFromHome <-as.character(data_parents$DistanceFromHome)
data_parents$DistanceFromHome[data_parents$DistanceFromHome== "[0,5]"] <- "close" 
data_parents$DistanceFromHome[data_parents$DistanceFromHome== "(5,15]" ] <- "far"
data_parents$DistanceFromHome[data_parents$DistanceFromHome == "(15,30]"] <- "far"
data_parents$DistanceFromHome <-as.factor(data_parents$DistanceFromHome)

```


```{r}
plot(pc.stable(data_parents) )
```

Let's try again with the whole dataset simplifying the variables more:
```{r}
names <- c(colnames(data))
for (var in names) {
  print(var)
  print(levels(data[,var]))
}
```


```{r}
data$WorkLifeBalance <-as.character(data$WorkLifeBalance)
data$WorkLifeBalance[data$WorkLifeBalance== 1 ] <- "low"
data$WorkLifeBalance[data$WorkLifeBalance== 2] <- "low"
data$WorkLifeBalance[data$WorkLifeBalance == 3] <- "high"
data$WorkLifeBalance[data$WorkLifeBalance == 4] <- "high"
data$WorkLifeBalance <-as.factor(data$WorkLifeBalance)

```

```{r}
data$TotalWorkingYears <-as.character(data$TotalWorkingYears)
data$TotalWorkingYears[data$TotalWorkingYears== "[0,5]" ] <- "less than 5 or equal"
data$TotalWorkingYears[data$TotalWorkingYears== "(5,10]"] <- "more than 5"
data$TotalWorkingYears[data$TotalWorkingYears == "(10,20]"] <- "more than 5"
data$TotalWorkingYears[data$TotalWorkingYears == "(20,40]"] <- "more than 5"
data$TotalWorkingYears <-as.factor(data$TotalWorkingYears)

```

```{r}
data$YearsAtCompany <-as.character(data$YearsAtCompany)
data$YearsAtCompany[data$YearsAtCompany== "[0,5]" ] <- "less than 5 or equal"
data$YearsAtCompany[data$YearsAtCompany== "(5,10]"] <- "more than 5"
data$YearsAtCompany[data$YearsAtCompany == "(10,20]"] <- "more than 5"
data$YearsAtCompany[data$YearsAtCompany == "(20,40]"] <- "more than 5"
data$YearsAtCompany <-as.factor(data$YearsAtCompany)

```

```{r}
data$NumCompaniesWorked <-as.character(data$NumCompaniesWorked)
data$NumCompaniesWorked[data$NumCompaniesWorked== "[0,1]"] <- "0 or 1"
data$NumCompaniesWorked[data$NumCompaniesWorked== "(1,4]"] <- "more than 1"
data$NumCompaniesWorked[data$NumCompaniesWorked == "(4,9]"] <- "more than 1"
data$NumCompaniesWorked <-as.factor(data$NumCompaniesWorked)

```

```{r}
data$NumCompaniesWorked <-as.character(data$NumCompaniesWorked)
data$NumCompaniesWorked[data$NumCompaniesWorked== "[0,1]"] <- "0 or 1"
data$NumCompaniesWorked[data$NumCompaniesWorked== "(1,4]"] <- "more than 1"
data$NumCompaniesWorked[data$NumCompaniesWorked == "(4,9]"] <- "more than 1"
data$NumCompaniesWorked <-as.factor(data$NumCompaniesWorked)

```

```{r}
data$Age <-as.character(data$Age)
data$Age[data$Age== "[17,29]"] <- "young"
data$Age[data$Age== "(29,40]"] <- "middle-age or old"
data$Age[data$Age == "(40,50]"] <- "middle-age or old"
data$Age[data$Age == "(50,60]"] <- "middle-age or old"
data$Age <-as.factor(data$Age)

```

```{r}
data$TotalSatisfaction <-as.character(data$TotalSatisfaction)
data$TotalSatisfaction[data$TotalSatisfaction== 1] <- "low"
data$TotalSatisfaction[data$TotalSatisfaction== 2] <- "low"
data$TotalSatisfaction[data$TotalSatisfaction == 3] <- "high"
data$TotalSatisfaction[data$TotalSatisfaction == 4] <- "high"
data$TotalSatisfaction <-as.factor(data$TotalSatisfaction)

```

```{r}
data$MonthlyIncome <-as.character(data$MonthlyIncome)
data$MonthlyIncome[data$MonthlyIncome== "[0,3e+03]"] <- "low"
data$MonthlyIncome[data$MonthlyIncome== "(3e+03,8e+03]"] <- "medium"
data$MonthlyIncome[data$MonthlyIncome == "(8e+03,1.2e+04]"] <- "medium"
data$MonthlyIncome[data$MonthlyIncome == "(1.2e+04,2e+04]"] <- "high"
data$MonthlyIncome <-as.factor(data$MonthlyIncome)

```

```{r}
data$StockOptionLevel <-as.character(data$StockOptionLevel)
data$StockOptionLevel[data$StockOptionLevel== 0] <- "No stock"
data$StockOptionLevel[data$StockOptionLevel== 1] <- "stock"
data$StockOptionLevel[data$StockOptionLevel == 2] <- "stock"
data$StockOptionLevel[data$StockOptionLevel == 3] <- "stock"
data$StockOptionLevel <-as.factor(data$StockOptionLevel)

```

```{r}
data$BusinessTravel <-as.character(data$BusinessTravel)
data$BusinessTravel[data$BusinessTravel== "Non-Travel"] <- "Non-Travel" 
data$BusinessTravel[data$BusinessTravel== "Travel_Rarely" ] <- "Travel"
data$BusinessTravel[data$BusinessTravel == "Travel_Frequently"] <- "Travel"
data$BusinessTravel <-as.factor(data$BusinessTravel)

```

```{r}
data$DistanceFromHome <-as.character(data$DistanceFromHome)
data$DistanceFromHome[data$DistanceFromHome== "[0,5]"] <- "close" 
data$DistanceFromHome[data$DistanceFromHome== "(5,15]" ] <- "far"
data$DistanceFromHome[data$DistanceFromHome == "(15,30]"] <- "far"
data$DistanceFromHome <-as.factor(data$DistanceFromHome)

```


```{r}
net_pc <- pc.stable(data)
plot(net_pc)
```

Let's see each variable with its own levels:
```{r}
names <- c(colnames(data))
for (var in names) {
  cat("\n")
  print(var)
  print(levels(data[,var]))
}
```


###############################################################################################################
TABU SEARCH
###############################################################################################################

```{r}
net_tabu <- tabu(data,score="bic") 
plot(net_tabu)
```


########################################################################################

Comparing the distances between the networks, through the hamming distance

```{r}

graph <- graphLayout(dagitty('dag {
{
Age -> Attrition
Age -> Education
Age -> JobLevel
Age -> MaritalStatus
Age -> NumCompaniesWorked
Age -> TotalWorkingYears
Age -> YearsAtCompany
BusinessTravel -> Attrition
DistanceFromHome -> Attrition
DistanceFromHome -> WorkLifeBalance
Education -> JobLevel
Gender -> BusinessTravel
JobLevel -> MonthlyIncome
MaritalStatus -> Attrition
MonthlyIncome -> Attrition
NumCompaniesWorked -> YearsAtCompany
OverTime -> Attrition
OverTime -> TotalSatisfaction
PercentSalaryHike -> Attrition
StockOptionLevel -> Attrition
StockOptionLevel -> MonthlyIncome
TotalSatisfaction -> Attrition
TotalWorkingYears -> JobLevel
TotalWorkingYears -> MonthlyIncome
TotalWorkingYears -> NumCompaniesWorked
TotalWorkingYears -> YearsAtCompany
WorkLifeBalance -> TotalSatisfaction
YearsAtCompany -> JobLevel
YearsAtCompany -> YearsSinceLastPromotion
YearsSinceLastPromotion -> Attrition
}
}
'))


plot(graph)
```

Let's see if the DAGs are different

all.equal tells us only if the DAGs are different.

```{r}
all.equal(net_pc, net_tabu)
```

```{r}
net_original <-model2network(toString(graph,"bnlearn"))

all.equal(net_pc, net_original)
```

```{r}

all.equal(net_tabu, net_original)
```

Computing the Hamming distance between DAGs
The Hamming distance gives the number of different edges, taking into account also their orientations.

PC vc Tabu
```{r}
hamming(net_pc, net_tabu)
shd(net_pc, net_tabu)
```

PC vc original
```{r}
hamming(net_pc, net_original)
shd(net_pc, net_original)
```

original vc Tabu
```{r}
hamming(net_original, net_tabu)
shd(net_original, net_tabu)
```


The function compare is used to compare the skeletons of the two networks, ignoring arc directions.
True positive (tp) arcs, which appear both in target and in current.
False positive (fp) arcs, which appear in current but not in target.
False negative (fn) arcs, which appear in target but not in current.

```{r}
compare(net_pc,net_tabu)
```

```{r}
compare(net_original,net_pc)
```

```{r}
compare(net_tabu,net_original)
```




Graphical comparison:
False positive arcs (which are missing or have different directions in the true network) are in red;
False negative arcs are in blue, and drawn using a dashed line.
(see documentation at http://www.bnlearn.com/examples/compare-dags/)
```{r}
 graphviz.compare(net_tabu,net_pc,net_original)
```

Our difference measures are anyway influenced by the fact that the pc-stable algorithm is able to detect less edges than the others.


####################################################################################################################

How do the results change varying one parameter per algorithm?

alpha = 0.05 (default)
```{r}
plot(net_pc)
```

```{r}
plot( pc.stable(data,alpha=0.20))
```

```{r}
plot( pc.stable(data,alpha=0.5))
```

```{r}
plot( pc.stable(data,alpha=0.75))
```

```{r}
bn <-  pc.stable(data,alpha=0.20)
cat("\n")
hamming(net_pc,bn)
shd(net_pc, bn)
```

```{r}
bn <-  pc.stable(data,alpha=0.5)
cat("\n")
hamming(net_pc,bn)
shd(net_pc, bn)
```

```{r}
bn <-  pc.stable(data,alpha=0.75)
cat("\n")
hamming(net_pc,bn)
shd(net_pc, bn)
```



Tabu Search

tabu = 10 (default). tabu also determine the value for max.tabu
```{r}
plot(tabu(data,score="bic"))
```

```{r}
plot(tabu(data,score="bic",tabu = 20))
```

```{r}
plot(tabu(data,score="bic",tabu = 50))
```

```{r}
plot(tabu(data,score="bic",tabu = 100))
```

Let's now see for lower values of tabu
```{r}
plot(tabu(data,score="bic",tabu = 5))
```

```{r}
plot(tabu(data,score="bic",tabu = 1))
```

```{r}
bn <-  tabu(data,score="bic",tabu = 20)
cat("\n")
hamming(net_tabu,bn)
shd(net_tabu, bn)
```

```{r}
bn <-  tabu(data,score="bic",tabu = 50)
cat("\n")
hamming(net_tabu,bn)
shd(net_tabu, bn)
```

```{r}
bn <-  tabu(data,score="bic",tabu = 100)
cat("\n")
hamming(net_tabu,bn)
shd(net_tabu, bn)
```

```{r}
bn <-  tabu(data,score="bic",tabu = 5)
cat("\n")
hamming(net_tabu,bn)
shd(net_tabu, bn)
```

```{r}
bn <-  tabu(data,score="bic",tabu = 1)
cat("\n")
hamming(net_tabu,bn)
shd(net_tabu, bn)
```

It seems that only the orientation of some adges changes by varying the value of tabu.